---
title: "Dealing with duplicate rows in R"
author: "Shefali C."
date: "2023-08-17"
output: 
        html_document:
          toc: TRUE
          theme: united
          highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This notebook contains different ways to find/remove duplicate rows from a dataframe.  


```{r load-libraries, warning=FALSE, message=FALSE}

library(dplyr)

```

#### 1. Removing rows through indexing. 

```{r create-df1}
#create dataframe
(d1 <- data.frame(name = c("Harry", "Ron", "Hermoine", "Fred", "George"),
                 value = c(10.10, 22.9, 56.7, 78.5, 32.8)
                 )
 )
#View(d1)
```

Simply use the index of the row to be deleted.  

```{r del-1row}

#delete 3rd row
(d2 <- d1[-3,])

```

```{r del-multiple-rows}
(d3 <- d1[-c(2,3),])

```

Taking example of iris dataset...

A little about the data: 

1. It contains info about 3 species of **iris** flower- ***setosa, versicolor, virginica.***  
2. The petal and sepal measurements along with width of each flower sample is present (in centimeters).  


```{r}
d4 <- dplyr::as_tibble(iris)
```


#### 2. Using **`duplicated()`**  

This function returns TRUE or FALSE for repeated occurance of a value.  
It will return FALSE for first occurrence; for subsequent occurrences of the same value, TRUE is returned indicating presence of a duplicate.  

This check for duplicates done below runs very, very slow for large datasets (~ thousands of rows).  

```{r}
#check whether any duplicate row present- returns TRUE or FALSE
any(duplicated(d4))
``` 

Removing the duplicate rows:  

```{r using-duplicated}
#check number of rows before deletion
print(paste0("Total rows before deletion of duplicates from iris: ", nrow(d4)))

#remove duplicate
d5 <- d4[!duplicated(d4),]

#number of rows after deletion
print(paste0("Number of rows after deletion: ", nrow(d5)))
```

#### 3. Using **`unique()`**  

**`unique()`** does the opposite of **`duplicated()`**. It retains only unique rows and removes duplicates.  
It is a part of base R.

```{r using-unique}
d6 <- unique(d4)
print(paste0("Number of rows after deletion: ", nrow(d6)))

#Another way
# d6 <- d4 %>% unique()
```

#### 4. Using **`distinct()`**  

**`distinct()`** does the same work as `unique()`. It is a part of **dplyr** package. It is a more efficient version of **`unique()`**.  

```{r using-distinct}
d7 <- distinct(d4)
print(paste0("Number of rows after deletion: ", nrow(d4)))

#Another way
# d7 <- d4 %>% distinct()

```

```{r}

```




