---
title: "Statistical Analysis of mtcars dataset"
author: "Shefali C."
date: "2024-02-05"
output: 
      prettydoc::html_pretty:
        theme: cayman
        highlight: vignette
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This notebook includes methods to visualize data distribution using ggplot2. I've used the built-in dataset **`mtcars`** for demonstration.  
Following plots and their customizations are included in this notebook:  
1. Histograms  
2. Density Plots  
3. Boxplots  

Other statistical measurements like mean, median, mode, standard deviation, variance have also been plotted towards the end of this notebook.  

So, all in all, I've taken the example of **`mtcars`** to show how a dataset is statistically analysed in initial phases of analysis.  

```{r load-libraries, message=FALSE, warning=FALSE}
#load libraries

library(tidyverse)
#textGrob() of grid package has been used.
library(grid)
``` 

```{r read-data}
#read data
df1 <- mtcars
```


