---
title: 'ggplot2: Usage of scale_(x|y)_* functions'
author: "Shefali C."
date: "2024-02-17"
output: 
    html_document:
      toc: true
      toc_float: 
        collapsed: false
        smooth_scroll: false
      theme: cosmo
      highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.width=4, fig.height=4, fig.align='center')
```

This notebook explains the **`expand`** argument of **`scale_(x|y)_continuous()`** and **`scale_(x|y)_discrete()`** functions.  
Both these functions are used to add/remove space between the main graph and X-Y axis.  

```{r load-library, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
#to use font of choice
library(showtext)
#to use emoji
library(emo)
#to stitch some graphs together
library(patchwork)
```

I like this font ***'Roboto Mono'*** from Google fonts, hence adding it to my graphs `r emo::ji("smile")`.

```{r add-font, warning=FALSE}
font_add_google('Roboto Mono', 'Roboto Mono')
#turn on automatic usage of showtext pacakge & 
#use the downloaded font for all plots, until changed.
showtext::showtext_auto()
```


Here's a simple bar chart without using **`scale_y_continuous()`**.

```{r plot1}
(base_plot <- diamonds %>% 
              dplyr::group_by(cut) %>% 
              dplyr::summarise(total_count = n()) %>% 
              ggplot(aes(x=cut, y = total_count))+
              geom_bar(stat = "identity", color = "black", fill = "#ffa600")+
              labs(title = "Total number of diamonds of each type of cut",
                   y = "total count in dataset", x="")+
              theme_bw()+
              theme(axis.text = element_text(face = "bold", color = "black", size = 12),
                    axis.title = element_text(size = 11),
                    plot.title = element_text(face = "bold", size = 16)))
```

In order to remove the space between the bars and X-axis, we can use **`scale_y_continuous()`** with its **"expand"** argument. 

## **1. Expand = c(a,b)**

1. Contains 2 arguments. ('a' and 'b' are just placeholders for easier  explanation)  
2. **argument 1 'a':**  
    - multiplicative factor.  
    - Works together with the 'limits' argument.  
    - increases or decreases space from BOTH ends of the axis. For e.g. in **`scale_y_continuous()`**, it adds/removes space from both lower limit, (below the bars) and upper Y limit.  
            
3. **argument 2 'b':**  
    - additive factor.  
    - add the specified unit to the given multiplicative factor.  
    
4. Formula:  
        - lower_Y_limit - ((range of Y-axis)*a + b)  
        - upper_Y_limit + (range of Y-axis)*a + b
        
5. Example:  
        - **`scale_y_continuous(limits = (0,6), expand = c(2,0))`**  
            - new_lower_y_limit = 0 - (6-0)*2 - 0 = -12.  
            - new_upper_y_limit = 6 + (6-0)*2 + 0= 18.  
            So, instead of limits being (0,6), now limits will be (-12,18) on Y axis and space will be added at both ends. 

        
## **2. Using different values for `expand`:**
        
### **i) expand = c(0,0)**  

```{r plot2}
#adding scale_y_cont.
base_plot+
  scale_y_continuous(expand = c(0,0))
```

#### **Explanation of logic using chart above:**

- Above, **"limits"** argument not present in `scale_y_continuous(expand = c(0,0))`. So, limits are automatically set by ggplot2, here it is (0,22000). [Since max count is of ideal diamonds with 21,551 data points.] 

- The values are:  
  - lower_y_limit = 0,  
  - range of Y axis = (22000 - 0) = 22000,  
  - multiplication factor = 0,  
  Using the equation above:  
  - $units\_of\_space\_below = 0 - 0*0 = 0$, ***hence no space below bars***.  
  - $units\_of_\_space\_above = 22000 + 22000*0 = 0$, ***hence no space above and bar sticks to upper panel line***.  

### **ii) Modify the 2nd argument (additive factor)**  

```{r plot3}
base_plot + 
  scale_y_continuous(expand = c(0, 150))
```

Changing additive factor to 150 adds a little space below and above the bars.  
**NOTE: ** changing this value to small values like 2 or 10, doesn't really reflect any changes in plot.

### **iii) Modify the first argument (multiplicative factor)**

```{r plot4}
base_plot +
  scale_y_continuous(expand = c(2,0))
```

#### **Explanation:**  

- previous Y axis limits = (0,22000)  
- changes:  
    - $new\_lower\_limit = 0 - 22000*2 - 0 = -44000$, ***hence you see the negative Y axis extends a little below -40000***.  
    - $new\_upper\_limit = 22000 + 22000*2 + 0 = 66,000$, ***hence +ve Y axis extends beyond 60,000***.  
    
### **iv) Change both factors:**  

```{r plot5}
base_plot +
  scale_y_continuous(expand = c(1.5, 150))
```

#### **EXplanation:**  

- Y axis extends from (-33150, 55,150)  
    - lower y : $0 - 22000*1.5 + 150 = -33,150$,  
    - upper y : $22000 + 22000*1.5 + 150 = 55,150$  
    
### **v) limits = c(m,n)** with **expand = c(a,b)**: 

The calculations are same as above, the only difference is here, I've explicitly specified limits.  
**NOTE: **  
- Without expand, there's little space below bars and the above 25000 label on Y axis.  
- With expand, the space below gets omitted & space above 25,000 on Y axis is also removed. 


```{r plot6, fig.width=8}
p1 <- base_plot + 
        scale_y_continuous(limits = c(0,25000))+
        labs(subtitle = "scale_y_cont(limits = c(0,25000))")+
        theme(plot.subtitle = element_text(size = 13, face = "bold.italic"))
p2 <- base_plot +
        scale_y_continuous(limits = c(0,25000), expand = c(0,0))+
        labs(subtitle = "scale_y_cont(limits = c(0,25000), expand = c(0,0))")+
        theme(plot.subtitle = element_text(size = 13, face = "bold.italic"))

p1 + p2
```

- **DRAWBACK of using `expand=c(a,b)`**- space gets added or removed from both ends of axis.  
- **What if we only want to add/remove padding from one end?? ie. remove space below the bars but the top Y axis should not be affected...**  

- **ALTERNATIVE: ** **`expansion()`** function can be used to achieve this.  

### **vi) expand = expansion(mult = c(a,b), add = (m,n))**

- **mult:** multiplicative factor. 'a' is for lower Y axis or left X axis, 'b' is for higher Y axis or right-end of X axis.  
- **add:** additive factor; m is for lower end, n is for higher end.  

In the code chunk below, keeping **`expand = expansion(mult = c(0,0.05))`** implies no space below '0' mark on Y axis and add 5% space above '22,000' mark. 

So, upper y axis becomes:  
$22000 + 22000*0.05 = 22000 + 1100 = 23,100$

```{r plot7}
base_plot+
  scale_y_continuous(expand = expansion(mult = c(0,0.05)))
```

## **3. Using expansion() on the discrete X axis:** 

Below, the 'add' argument has been used for expansion in `scale_x_discrete()` function. It adds that unit of space on either ends of X-axis.

```{r plot8, fig.width=8}
p3 <- base_plot+
        scale_y_continuous(expand = expansion(mult = c(0,0.05)),
                           limits = c(0,23000),
                           breaks = seq(0,23000,4000),
                           labels = paste0(seq(0,23000,4000)/1000,"K"))+
        ##NOTE--USAGE OF add HERE
        scale_x_discrete(expand = expansion(add = c(1,1)))+
        labs(subtitle = "used: scale_x_discrete(expand = expansion(add = c(1,1)))")+
        theme(plot.subtitle = element_text(size = 13, face = "bold.italic"))

p4 <- base_plot+
        scale_y_continuous(expand = expansion(mult = c(0,0.05)),
                           limits = c(0,23000),
                           breaks = seq(0,23000,4000),
                           labels = paste0(seq(0,23000,4000)/1000,"K"))+
        ##NOTE--USAGE OF add HERE
        scale_x_discrete(expand = expansion(add = c(1,2)))+
        labs(subtitle = "used: scale_x_discrete(expand = expansion(add = c(1,2)))")+
  theme(axis.title.y = element_blank(),
        plot.subtitle = element_text(size = 13, face = "bold.italic"))

p3 + p4
```