---
title: "Colors in R"
author: "Shefali C."
date: "2023-09-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

A bunch of ways to access different colors/color palettes in R.  

[Notebook gets updated whenever I forget how to access different color palettes and feel a need to note down the techniques somewhere!]

```{r}
library(ggplot2)
library(scales)
library(colorspace)
```

### **`HCL`** palette:  
1. Refer [this](https://blog.r-project.org/2019/04/01/hcl-based-color-palettes-in-grdevices/) page for detailed reading.  

2. A better article for seeing color swatches in hcl package- [hcl-palettes](https://colorspace.r-forge.r-project.org/articles/hcl_palettes.html)  

3. Useful functions in **`colorspace`** package: [colorspace functions](https://colorspace.r-forge.r-project.org/articles/palette_visualization.html).


#### 1. List of all color palettes available in hcl package.  

```{r}
#list of all hcl palettes
hcl.pals()
```

#### 2. Function to visualize color swatches for Sequential palettes:  


```{r}
colorspace::hcl_palettes("Sequential (single-hue)", plot = T)
```

```{r}
colorspace::hcl_palettes(type = "Sequential (multi-hue)", plot = T)
```

```{r}
colorspace::hcl_palettes(type = "seq", plot = T)
```







